Conversion of FFCS ICE fleet to EV: a data-driven approach

### 2020-11-17 Chapter4/4_0_abstract.tex
chapter_4_passed
conclusion_added

\section{Abstract}
\label{sec:4_0_abstract}

The understanding of the mobility on urban spaces is useful for the creation of smarter and sustainable cities. However, getting data about urban mobility is challenging, since only a few companies have access to accurate and updated data, that is also privacy-sensitive.

In this work, we characterize three distinct car-sharing systems which operate in Vancouver (Canada) and nearby regions, gathering data for more than one year. The study uncovers patterns of users' habits and demands for these services. 
We highlight the common characteristics and the main differences among car-sharing systems. Finally, we believe our study and data is useful for generating realistic synthetic workloads.

This chapter refers mostly the paper \textit{Characterizing client usage patterns and service demand for car-sharing systems} \cite{VancouverCS}, published on the Journal, available online since October 11, 2019. My contribution is mainly focused in sections \mc{mettere qui le sezioni}
\mc{rivedere le cose I we e quali sezioni io ho effettivamente dato contributo}

>>>>

%\section{Abstract}
%\label{sec:4_0_abstract}

%The understanding of the mobility on urban spaces is useful for the creation of smarter and sustainable cities. However, getting data about urban mobility is challenging, since only a few companies have access to accurate and updated data, that is also privacy-sensitive. 
%
%In this work, we characterize three distinct car-sharing systems which operate in Vancouver (Canada) and nearby regions, gathering data for more than one year. The study uncovers patterns of users' habits and demands for these services. 
%We highlight the common characteristics and the main differences among car-sharing systems. Finally, we believe our study and data is useful for generating realistic synthetic workloads.

This chapter refers mostly the paper \textit{Characterizing client usage patterns and service demand for car-sharing systems} \cite{VancouverCS}, published on the Journal, available online since October 11, 2019. My contribution is mainly focused in all the aspects about car2go analyses.

### 2020-11-17 Chapter6/6_1_intro.tex
conclusion_added

After quickly discussing related work in section~\ref{sec:6_2_related}, I present and characterize data in section~\ref{sec:6_3_data}, and the simulation model and tool in section~\ref{sec:6_4_Modelling}. section~\ref{sec:6_6_freefloating} discusses the impact of charging stations placement policies, while section~\ref{sec:6_7_results} compares return policies. Section~\ref{sec:6_8_resPoles} presents the impact of concentrating or spreading charging stations in the city. Finally section~\ref{sec:6_9_conclusion} concludes the chapter.

>>>>

After quickly discussing related work in section~\ref{sec:6_2_related}, I present and characterize data in section~\ref{sec:6_3_data}, 
%and the simulation model and tool in section~\ref{sec:6_4_Modelling}. 
section~\ref{sec:6_6_freefloating} discusses the impact of charging stations placement policies, while section~\ref{sec:6_7_results} compares return policies. Section~\ref{sec:6_8_resPoles} presents the impact of concentrating or spreading charging stations in the city. Finally section~\ref{sec:6_9_conclusion} concludes the chapter.

### 2020-11-17 Chapter8/8_0_abstract.tex
chapt
conclusion_added

\section{Abstract}
Free Floating Car Sharing (FFCS) services are a flexible alternative to car ownership. These transportation services show highly dynamic usage both over different hours of the day, and across different city areas. 
In this work, we study the problem of predicting FFCS demand patterns -- a problem of great importance to an adequate provisioning of the service. We tackle both the prediction of the demand i) over time and ii) over space. 
We rely on months of real FFCS rides in Vancouver, which constitute our ground truth. We enrich this data with detailed socio-demographic information obtained from large open-data repositories to predict usage patterns. 
Our aim is to offer a thorough comparison of several machine learning algorithms in terms of accuracy and easiness of training, and to assess the effectiveness of current state-of-art approaches to address the prediction problem.
Our results show that it is possible to predict the future usage with relative errors down to 10\%, and the spatial prediction can be estimated with relative errors of about 40\%.
Our study also uncovered the socio-demographic features that most strongly correlate with FFCS usage, providing interesting insights for providers interested into opening service in new regions.

@@

\mc{in questo capitolo tengo il we tranne nella caratterizzazione degli open data e nelle spatial analyses}

>>>>

%\section{Abstract}
%Free Floating Car Sharing (FFCS) services are a flexible alternative to car ownership. These transportation services show highly dynamic usage both over different hours of the day, and across different city areas. 
%In this work, we study the problem of predicting FFCS demand patterns -- a problem of great importance to an adequate provisioning of the service. We tackle both the prediction of the demand i) over time and ii) over space. 
%We rely on months of real FFCS rides in Vancouver, which constitute our ground truth. We enrich this data with detailed socio-demographic information obtained from large open-data repositories to predict usage patterns. 
%Our aim is to offer a thorough comparison of several machine learning algorithms in terms of accuracy and easiness of training, and to assess the effectiveness of current state-of-art approaches to address the prediction problem.
%Our results show that it is possible to predict the future usage with relative errors down to 10\%, and the spatial prediction can be estimated with relative errors of about 40\%.
%Our study also uncovered the socio-demographic features that most strongly correlate with FFCS usage, providing interesting insights for providers interested into opening service in new regions.

@@

### 2020-11-17 thesis.tex
conclusion_added

\summary%\sommario

This is where you write your abstract \dots\ (Maximum 4000 characters, i.e. maximum two pages in normal sized font, typeset with the thesis layout).

The abstract environment is also available, but  \texttt{\string\summary} is preferred because it generates an un-numbered chapter. The abstract environment is more suitable for articles and two column typesetting without a separate title page.

\emptypage %  it works even without the classica option
\acknowledgements% or \ringraziamenti

And I would like to acknowledge \dots

Acknowledgements are mandatory only when people outside the academic institution supported the development of the research that was performed in order to reach the conclusion of the doctorate program.

>>>>

\summary%\sommario
%This is where you write your abstract \dots\ (Maximum 4000 characters, i.e. maximum two pages in normal sized font, typeset with the thesis layout).
%The abstract environment is also available, but  \texttt{\string\summary} is preferred because it generates an un-numbered chapter. The abstract environment is more suitable for articles and two column typesetting without a separate title page.
\mc{intro}

In my thesis, I presented a methodology to address, in different cases of studies, all the challenges related to the conversion of combustion engine cars to electric vehicles in FFCS. In particular, the main drivers focus on a profitable and technically sustainable system setup able to guarantee a flexible and appealing mobility service to an increasing customer audience.

In the first part of my thesis, I described the software I developed in order to scrape from the web real combustion engine FFCS, from two providers: car2go and Enjoy. The data collection car2go data collection lasted from \mc{december 2016 to january 2018, collecting more than 28 million users' bookings spread in 25 cities} while the Enjoy data collection phase started \mc{may 2018 and lasted until BOH}, collectin about \mc{BOH} bookings in 6 cities.

Then, I characterized both datasets in Turin, one of the cities in which both FFCS providers work. Initially, I defined a filtering criterion to remove outliers from the dataset and identify only real users' ride: entries reporting a trip where the user realistically moved. These analyses allowed me to identify geo-spatial user's patterns. After that, I compared the car2go customer's pattern with the one-way and two-way car-sharing system. The results showed how users prefer more flexible services like FFCS or one-way car sharing.

Once the data are consolidated, I developed (i) a methodology to assign a charging station to a city by looking at users' pattern; (ii) system policies to manage the fleet when the vehicle SoC may not guarantee a trip; (iii) an event-based trace-driven simulator able to replicate the recorded trips in an electrified scenario evaluating the feasibility of each configuration.

At this point I studied in Berlin, Milan, Turin, and Vancouver (cities having different traffic patterns) different electric FFCS setup. The best logic to place a charging station is to cover the zones recording more parkings: increasing the charging station presence in those zones, the probability to find a free plug increases too. Then, I showed how an electric FFCS is sustainable with few charging stations (8-10~\% of city zones) if the users accept to park the car in a charging station when their ride finishes nearby or when the SoC goes below a security threshold. Finally, I investigate different charging station setups. The best results come by balancing the spread of charging stations: concentrated all plugs in a unique hub or spread them all above the city leads to poor performances in terms of system sustainability and user's discomforts.

To still reduce the number of charging stations to have a sustainable electric FFCS, I compared several optimization algorithms. My results showed how a Genetic Algorithm is able to find a better solution able to shrink the minimum amount of resources to sustain the same mobility demand.

After that, I moved my attention to the users' rentals predictability. The main goal is to understand how different open-data sources could impact on the recorded FFCS users' rental. I used as a case of study the city of Vancouver, augmenting the rentals dataset with the weather and socio-economic open data. Initially, I compared several time-series forecasts to predict the users' demand in the short and medium-term. The Random Forrest regression produced better accuracy and results interpretability. Then I correlated the socio-economics features characterizing each Vancouver neighborhood to FFCS demand and again, the Random Forrest regression clearly showed a subset of socio-economic features more correlated to the FFCS users' rentals.

Finally, I questioned the system scalability figuring out several scenarios having increasing demand. I used a model able to synthesize users' demand by looking only at the geo-spatial users' rentals. By varying the electric FFCS setup and simulating the new scenario I pointed out how a linear increase in the demand intensity leads to a fleet sub-linear increase. Finally, I projected those considerations in euros proofing how electric FFCS has rooms for economic growth.
